Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm

### 2011-08-25 ltx/data_paper/DataPaper.tex
edits to discussion

\section{Discussion}\label{sec:EM-Discussion}
This paper is an extension of previous work in developing a framework for subject-specific neural field modelling. The extensions of the previous work include a method for estimating the support of the connectivity kernel and disturbance covariance, and the development of an EM framework for the neural field equations using the URTSS. The EM framework combined with the estimation of the support of the disturbance enables estimation of the temporal disturbance and noise characteristics of the system. Furthermore, by estimating the support of the connectivity kernel we are able to relax the strong assumption of isotropy in the estimated connectivity structure.

@@

An alternative to estimating the support of the connectivity kernel is to place a regular grid of connectivity kernel basis function over the entire spatial domain of the field. However, the estimate of the spatial extent of the connectivity kernel greatly improves the speed of the estimation algorithm as the number of basis functions for the kernel decomposition can be significantly reduced. This also improves the uncertainty in the estimated coefficients, as a lower number of unknown parameters are estimated. The approximated support of the disturbance covariance function facilitates the application of the EM algorithm. This way the disturbance covariance matrix can be decomposed into two parts: an unknown scalar and a constant matrix depending only on the inferred spatial support. Therefore, the estimation problem of the disturbance covariance matrix breaks down to the estimation of a single scalar parameter, improving the accuracy and the convergence property of the algorithm.

\subsection{Development of the EM for the URTSS}
The additional Taylor approximation step introduced in the E-step of the algorithm resembles the extended Kalman filtering \cite{Jazwinski1970,Chui2009}. The EKF approximates a stationary non-linear dynamic system with a linear non-stationary one using the first order Taylor expansion around the state estimates. This might imply that the extended Kalman smoother (EKS) is a better candidate --- in terms of implementation --- for the state inference in the proposed algorithm. However, the EKS requires re-linearising about the backward estimates which in turn necessitates the non-linear mapping to be uniquely invertible. Although this is indeed the case in the neural field equation employed in this work it will impose more computational complexity in the state estimation. In addition, the URTSS provides better state estimations compared to the EKS \cite{Haykin2001} and therefore the constructed lower bound will be more accurate, resulting in a better parameter estimation.

While the EKS has been  applied in the EM algorithm for learning non-linear models \cite{Ghahramani1999} this is the first work (to the best author's knowledge) where the URTSS is used in this context. However, as in the case of the EKS the stability of the algorithm cannot be ensured. This is because the use of the EKS or URTSS represents an approximation to the exact E-step. As a result the maximisation of the approximated lower bound might not necessarily increase the log-likelihood function. Nevertheless, it is envisaged that the use of the URTSS will provide a more stable algorithm due to its superior performance.

>>>>

\section{Discussion}\label{sec:EM-Discussion}
This paper is an extension of previous work in developing a framework for subject-specific neural field modelling. The extensions of the previous work include a method for estimating the support of the connectivity kernel and disturbance covariance, and the development of an EM framework for the neural field equations using the URTSS. The EM framework combined with the estimation of the support of the disturbance enables estimation of the temporal disturbance and noise characteristics of the system. Furthermore, by estimating the support of the connectivity kernel we are able to relax the strong assumption of isotropy in the estimated connectivity structure.

@@

An alternative to estimating the support of the kernel is to place a regular grid of connectivity kernel basis function over the entire spatial domain of the field. However, the estimate of the spatial extent of the connectivity kernel greatly improves the speed of the estimation algorithm as the number of basis functions can be significantly reduced. This also improves the uncertainty in the estimated coefficients, as a lower number of unknown parameters are estimated. The approximated support of the disturbance covariance function facilitates the application of the EM algorithm for estimating the temporal variance in the disturbance signal. This way the disturbance covariance matrix can be decomposed into two parts: an unknown scalar and a constant matrix depending only on the inferred spatial support. Therefore, the estimation problem of the disturbance covariance matrix breaks down to the estimation of a single scalar parameter, improving the accuracy and the convergence property of the algorithm.

\subsection{Development of the EM for the URTSS}
The additional Taylor approximation step introduced in the E-step of the algorithm resembles the extended Kalman filter (EKF) \cite{Jazwinski1970,Chui2009}. The EKF approximates a stationary nonlinear dynamical system with a linear nonstationary one using the first-order Taylor expansion around the state estimates. This might imply that the extended Kalman smoother (EKS) is a better candidate --- in terms of implementation --- for the state inference in the proposed algorithm. However, the EKS requires re-linearising about the backward estimates, which in turn necessitates the nonlinear mapping to be uniquely invertible. Although this is indeed the case in the neural field equation employed in this work it will impose more computational complexity in the state estimation. In addition, the URTSS provides better state estimations compared to the EKS \cite{Haykin2001} and therefore the constructed lower bound will be more accurate, resulting in a better parameter estimation.

While the EKS has been applied in the EM algorithm for learning nonlinear models \cite{Ghahramani1999}, this is the first work (to the best author's knowledge) where the URTSS is used in this context. However, as in the case of the EKS, the stability of the algorithm cannot be ensured. This is because the EKS or URTSS provides an approximation to the exact E-step. Following this, the maximisation of the approximated lower bound is not guaranteed to increase the log-likelihood function. Nevertheless, it is envisaged that the use of the URTSS will provide a more stable algorithm due to its superior performance.
